### 2016-05-18 mcmc.tex
makefile builds figures; nicer figures

\begin{problem}\label{prob:MH}
In your scientific programming language of choice, write a very simple
M-H \MCMC\ sampler.
Sample in a single parameter $x$ and give the sampler as its density
function $p(x)$ a Gaussian density with mean 2 and variance 2.
(Note that variance is the \emph{square} of the standard deviation.)
Give the sampler a proposal distribution $q(x'\given x)$ a Gaussian
pdf with mean zero and variance 1.
Initialize the sampler with $x=0$ and run the sampler for more than $10^4$ steps.
Plot the results as a histogram, with the true density over-plotted sensibly.
\end{problem}

\begin{problem}
Re-do \problemname~\ref{prob:MH} but now with an input density
that is uniform on $3<x<7$ and zero everywhere else.
What change did you have to make to the initialization, and why?
\end{problem}

\begin{problem}\label{prob:twod}
Re-do \problemname~\ref{prob:MH} but now with an input density
that is a function of two variables $(x, y)$.
For the density function use two different functions.
The first density function is a covariant two-dimensional Gaussian density with variance tensor
\begin{eqnarray}
V &=& \left[\begin{array}{cc} 2.0 & 1.2 \\ 1.2 & 2.0 \end{array}\right]
\quad.
\end{eqnarray}
The second density function is a rectangular top-hat function that is uniform on
the joint constraint $3<x<7$ and $1<y<9$ and zero everywhere else.
For the proposal distribution $q(x', y'\given x, y)$ a two-dimensional Gaussian
density with zero mean and variance tensor set to the two-dimensional
identity matrix.
Plot the two one-dimensional histograms and also a two-dimensional scatter
plot for each sampling.
\end{problem}

\begin{problem}
Re-do \problemname~\ref{prob:MH} but with different values for
the variance of the proposal distribution $q(x'\given x)$.
What happens when you go to very extreme values (like for instance
$10^{-8}$ or $10^6$)?
\end{problem}

>>>>

\begin{problem}\label{prob:MH}
In your scientific programming language of choice, write a very simple
M-H \MCMC\ sampler.
Sample in a single parameter $x$ and give the sampler as its density
function $p(x)$ a Gaussian density with mean 2 and variance 2.
(Note that variance is the \emph{square} of the standard deviation.)
Give the sampler a proposal distribution $q(x'\given x)$ a Gaussian
pdf with mean zero and variance 1.
Initialize the sampler with $x=0$ and run the sampler for more than $10^4$ steps.
Plot the results as a histogram, with the true density over-plotted sensibly.
The resulting  plot should look something like \figurename~\ref{fig:MH}.
\end{problem}

\begin{figure}[!htbp]
\begin{center}
\includegraphics[width=0.5\textwidth]{problems/p2}
\end{center}
\caption{Solution to \problemname~\ref{prob:MH}:
MCMC samples from a Gaussian (black) and the true distribution (blue).}
\label{fig:MH}
\end{figure}

\begin{problem}\label{prob:MH2}
Re-do \problemname~\ref{prob:MH} but now with an input density
that is uniform on $3<x<7$ and zero everywhere else.
The plot should look like \figurename~\ref{fig:MH2}.
What change did you have to make to the initialization, and why?
\end{problem}

\begin{figure}[!htbp]
\begin{center}
\includegraphics[width=0.5\textwidth]{problems/p3}
\end{center}
\caption{Solution to \problemname~\ref{prob:MH2}:
MCMC samples from a uniform distribution (black) and the true distribution
(blue).}
\label{fig:MH2}
\end{figure}

\begin{problem}\label{prob:twod}
Re-do \problemname~\ref{prob:MH} but now with an input density
that is a function of two variables $(x, y)$.
For the density function use two different functions.
\emph{(a)} The first density function is a covariant two-dimensional Gaussian
density with variance tensor
\begin{eqnarray}
V &=& \left[\begin{array}{cc} 2.0 & 1.2 \\ 1.2 & 2.0 \end{array}\right]
\quad.
\end{eqnarray}
\emph{(b)} The second density function is a rectangular top-hat function that
is uniform on the joint constraint $3<x<7$ and $1<y<9$ and zero everywhere else.
For the proposal distribution $q(x', y'\given x, y)$ a two-dimensional Gaussian
density with zero mean and variance tensor set to the two-dimensional
identity matrix.
Plot the two one-dimensional histograms and also a two-dimensional scatter
plot for each sampling.
\end{problem}

\begin{figure}[!htbp]
\begin{center}
\includegraphics[width=0.7\textwidth]{problems/p4a}
\end{center}
\caption{Solution to \problemname~\ref{prob:twod}a:
MCMC samples from a two-dimensional Gaussian (scatter plot) and the
one-dimensional marginalized distributions (histograms).}
\label{fig:twod}
\end{figure}

\begin{figure}[!htbp]
\begin{center}
\includegraphics[width=0.7\textwidth]{problems/p4b}
\end{center}
\caption{Solution to \problemname~\ref{prob:twod}b:
MCMC samples from a two-dimensional top hat distribution (scatter plot) and the
one-dimensional marginalized distributions (histograms).}
\label{fig:twod}
\end{figure}

\begin{problem}\label{prob:MH-sigma}
Re-do \problemname~\ref{prob:MH} but with different values for
the variance of the proposal distribution $q(x'\given x)$.
What happens when you go to very extreme values (like for instance
$10^{-8}$ or $10^6$)?
\end{problem}

\begin{figure}[!htbp]
\begin{center}
\includegraphics[width=\textwidth]{problems/p5}
\end{center}
\caption{Solution to \problemname~\ref{prob:MH-sigma}:
MCMC samples from a two-dimensional top hat distribution (scatter plot) and the
one-dimensional marginalized distributions (histograms).}
\label{fig:MH-sigma}
\end{figure}
